---
title: "Using the MHW-CI database to plot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the MHW-CI database to plot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mhwci)
```

This shows some basic commands to connect to the MHW-CI database, summarize data and plot. 


If you are developing this package,  editing scriptsor adding packages, then this necessary 

```{r echo=TRUE}
devtools::load_all()
```

Here is a script that you can copy/paste into the console.  

This assumes you have set the variable MHWDBFILE either in the `.Renviron` file or in the environment/shell before starting R/Rstudio.    

MHWDBFILE must be the full path to the database file.  You can set this manually in your scripts is prefered


```
# read from the environment
db_file <- get_dbfile()
```

OR 

```
# set manually
db_file = "/full/path/to/mhwci.b"
```

```
# Optional check that the `db_file` is a file that can be found from your script. 

file.exists(db_file))

# Get connected
conn<- mhw_connect(db_file)
# optionally just showing that we have an actual connect
print(duckdb::dbListTables(conn))


# Each scenario has it's own table of marine heat wave metrics, so let's set that table name here
mhw_table <-"arise15_metrics"
```

You can run SQL command to get data using the duckdb database engine, which is very fast.  

This SQL averages duration for the specific table 

```{r echo=TRUE}
# this is a function that creates a string of SQL with the table we want to use
sql<- avg_duration_by_decade_sql(mhw_table)
print(sql)
```


Here is how you run a query: 

```{r echo=TRUE}
# run this SQL using the `dbGetQuery` command.  
duration_by_loc<- DBI::dbGetQuery(conn, sql)

head(duration_by_loc)
```

this does all the stuff above encapsulated into a function: 

```{r echo=TRUE}
duration_rasters <- durations_by_decade_raster(conn, mhw_table)
```

create a map of these rasters that manages the outliers

```{r echo=TRUE}
cut_percent = 0.5
duration_title <- paste("Mean MHW Duration (days) by Decade of Onset, ", mhw_table)

plot_rasters_squish_outliers(raster_list = duration_rasters, title = duration_title, cut_percent = 0.5 )
```
