---
title: "Using the MHW-CI database to plot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the MHW-CI database to plot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::clean_cache(clean = TRUE)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(magrittr)
library(mhwci)

```

This shows some basic commands to connect to the MHW-CI database, summarize data and plot. 


If you are developing this package,  editing scripts or adding packages, then run this in the console prior to this script: `devtools::load_all()`

This package requires a database file that stores all the data. You have to create a database 'connection' first

You can type the filename in a variable 

```{r, echo=TRUE}
# read from the environment
db_file <- "db/mhwci_v3.db"
```

or you can set the path to the database file in an  `.Renviron` file like this

`MHWDBFILE=/Users/billspat/Code/SpaCELab/marine_heatwave_work/mhwci/db/mhwci_v3.db`

Note that MHWDBFILE must be the full path to the database file.  If you do set it in the `.Renviron` file, then you can use this function:

```{r, echo=TRUE}
# read from the environment
db_file <- mhwci::get_dbfile()
```

Then you need to create a connection object: 

```{r, echo=TRUE}
db<- mhwci::mhw_connect(db_file)
print(db)
```

This isn't actually the database, just a connection to it required for all database functions. 

Now you could list the tables in the database:

```{r, echo=TRUE}
print(duckdb::dbListTables(db))
```

The database has a table for each scenario, listed above, and a few other support tables.  This package has a quick function to look at the first few rows of a table, similar to the Unix `head` command.  

```{r, echo=TRUE}

mhwci::table_head(db,"ssp245_decade_metrics")

```
## Calculating Metrics

Set the scenario table to use for this plot. 

```{r, echo=TRUE}
mhw_table <-"arise10_decade_metrics"
```

This function queries this table for average durations by point and decade into data frames with x, y and value, and then uses the `terra` package to create the rasters of those decades in a list.  It also rotates the world plots so that the center is along the prime meridian instead of the dateline. 

```{r, echo=TRUE}
raster_list <- durations_by_decade_raster(db, mhw_table)
print(raster_list)
```

Plot this list of rasters, setting the name of the metric (average duration) and the threshold of 1/2 year

```{r, echo=TRUE}
plot_rasters_squish_outliers(raster_list, title = mhw_table, 
                             scale_label = "Average Duration, d", 
                             )

```


## Plot duration other scenarios

Combine the commands above to plot the other scenarios.  Note we change the title manually.  This re-uses the connection object `db` we created above.   The `duration_rasters()` function calls the function to create and run the SQL on the database

```{r, echo=TRUE}
raster_list <- durations_by_decade_raster(db, mhw_table= "arise15_decade_metrics")
plot_rasters_squish_outliers(raster_list, title = "ARISE 1.5 Average Duration", 
                             scale_label = "Mean Duration, d", 
                             threshold = 365/2)


```
```{r, echo=TRUE}
raster_list <- durations_by_decade_raster(db, mhw_table= "ssp245_decade_metrics")
plot_rasters_squish_outliers(raster_list, title = "SSP2 4.5 Average Duration", 
                             scale_label = "Mean Duration, d", 
                             threshold = 365/2)


```

## Plot other metrics and summaries

There are functions to summary any of the metrics with various statistics (mean, median, min, max, mode, etc)

Let's take the average of the average intensity

```{r, echo=TRUE}
# set the table
mhw_table <- "arise10_decade_metrics"
# select the metric or column in the table above
mhw_metric <- 'int_mean'
# select the summary function, must be one of the SQL functions, not an R function
sql_function <- 'avg'

#create the rasters for each decade for this metric, similar to duration but tis is a generic function
raster_list <- metrics_by_decade_raster(db, mhw_table, mhw_metric = mhw_metric, sql_function = sql_function)

# plot, constructing the title from the variables we set above
# no threshold is set
plot_rasters_squish_outliers(raster_list, 
                             title = paste(mhw_table, mhw_metric, sql_function), 
                             scale_label = "avg intensity average, c" )
```


## Filtering ensembles

If you use a special syntax, you can select which ensembles to use in the summaries.  The format is a string with list of ensembles and leading zeros, like this `"006,007,008"`  Ten has a leading zero to, or `"010"`  This is not a vector but a string!  (for now).  See the code below

```{r, echo = TRUE}

ensemble_list_string = "006,007,008"

#create the rasters for each decade for this metric, similar to duration but tis is a generic function
raster_list <- metrics_by_decade_raster(db, mhw_table, mhw_metric = mhw_metric, sql_function = sql_function, 
                                        ensemble_list_string = ensemble_list_string )

# plot, constructing the title from the variables we set above
# no threshold is set
plot_rasters_squish_outliers(raster_list, 
                             title = paste(mhw_table, mhw_metric, sql_function, " ensembles ", ensemble_list_string), 
                             scale_label = "avg intensity average, c")
                             
```


## Using SQL directly 

You can run SQL command to get data using the duckdb database engine, which is very fast.  

This SQL averages duration for the specific table 

```{r, echo=TRUE}
# this is a function that creates a string of SQL with the table we want to use
sql<- avg_duration_by_decade_sql(mhw_table)
print(sql)
```


Here is how you run a query: 

```{r, echo=TRUE}
# run this SQL using the `dbGetQuery` command.  
duration_by_loc<- DBI::dbGetQuery(db, sql)
# show the first few rows
head(duration_by_loc)
```

Now close the connection used here, otherwise there may be errors ("could not set lock on file ...")


```{r, echo=TRUE}
DBI::dbDisconnect(db)
```