---
title: "MHW CI Exploratory Anlaysis"
output: md_document
date: "2024-09-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = FALSE)
```

## Marine Heat Wave Climate Intervention Modelling Project

Dr. Lala Kounta, Dr. Phoebe Zarnetske

### Database connection

First set the database file to use.  If there is on in the `.Renviron` file, this will retrieve that and check that it exists. 

```{r}
db_file <- mhwci::get_dbfile()
print(db_file)
print(file.exists(db_file))
```

If that is not set, set the db_file variable here and uncomment this line, but it may be easier to create the .Renviron file instead for your setup. 

```{r}
# db_file <- "/mnt/gs21/scratch/plz-lab/mhwci/db/mhwci.db"
# print(file.exists(db_file))
```

Now that the dbfile is set, create a connection that will be used for the remaining functions.  the variable name `conn` is convention but can be named anything.   

```{r echo=TRUE}

conn<- mhwci::mhw_connect(db_file)
print(conn)
```

### Database Contents
List of tables in current database:

```{r}
print(duckdb::dbListTables(conn))
```
*Sample of data in table arise10_metrics:*

Detection modelling metrics calculated for 10 ensembles under the Arise-1.0 scenario:

```{r echo=TRUE}
table_head(conn = conn, tablename = "arise10_metrics", n = 10)
```

```{r}
decades <- table_head(conn, 'decades')
print(decades)
```
### Decadal Analysis

There is a table of metrics (duration, dates, lat/long, intensity, etc) for each  scenario that contains all scenarios.  This sets the table to use. 

```{r echo=TRUE}
mhw_table <-"arise15_metrics"
```


*Histogram of durations by decade*

```{r echo=TRUE}
duration_by_decade_histogram(conn, mhw_table, log_scale = TRUE)
```

**Global heatwave duration by decade**

Note: 
we can improve on the title by selecting a single row from this table 
and getting the value in the 'scenario' column

```{r echo=TRUE}
duration_rasters <- durations_by_decade_raster(conn, mhw_table)
cut_percent = 0.5
duration_title <- paste("Mean MHW Duration (days) by Decade of Onset, ", mhw_table)

plot_rasters_squish_outliers(raster_list = duration_rasters, title = duration_title, cut_percent = 0.5 )

```

In this figure, any values above 99.5 percentile are combined or "squished" into the same color so that those huge outliers are still present but share the same color (red).  The legend still has to be corrected to include those high value or use the "break" method but that will take some work.  


### close database connection 

run this when finished so that other other connections to thie file (say from other parts of your code) can be made. 

```{r echo=TRUE}
print(duckdb::dbDisconnect(conn))

```


